### 2019-01-31 Example/Example.Rmd
Add example of problem

---
author:
  - name: FirstName LastName
    affiliation: University/Company
    address: >
      First line
      Second line
    email: \email{name@company.com}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "A Capitalized Title: Something about a Package \\pkg{foo}"
  # If you use tex in the formatted title, also supply version without
  plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
bibliography: index.bib   
output: rticles::jss_article
---

# Example

* This is unwanted output

[@R-sensitivity]

* This is what I wany to show in the paper

[@R-sensitivity]

# Referenes

```{r, include=FALSE}
knitr::write_bib('sensitivity', 'index.bib')
```