### 2019-01-31 html/manuscript.Rmd

---
author:
  - name: Yihui Xie
    affiliation: RStudio
    address: |
      First line  
      Second line
    email: <name@company.com>
    url: <https://rstudio.com>
  - name: Romain Lesur
    affiliation: Affiliation
title: "A Capitalized Title: Something about a Package **foo**"
shorttitle: "**foo**:  A Capitalized Title"
abstract: |
  The abstract of the article.
# at least one keyword must be supplied
keywords: [keywords, not capitalized, "[Java]{.proglang}"]
links-to-footnotes: true
paged-footnotes: true
bibliography: index.bib
output: 
  pagedown::jss_paged:
    self_contained: false
---

# Introduction

This template demonstrates some of the basic [markdown]{.proglang} you'll need to know to create a JSS article with **pagedown** [@R-pagedown].

## Code formatting

You should use **Pandoc**'s [markdown]{.proglang} but you also can use $\LaTeX$ commands:

* [Java]{.proglang} or $\proglang{Java}$ 
* **plyr**, [plyr]{.pkg} or $\pkg{plyr}$
* `print("abc")` or $\code{print("abc")}$

# [R]{.proglang} code

Can be inserted in regular [R Markdown]{.proglang} blocks.

```{r}
x <- 1:10
x
```

# Render the document

If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.

[@R-httk]

# References {-}

```{r, include=FALSE}
knitr::write_bib(c(.packages(), 'pagedown', 'xaringan', 'httk'), 'index.bib')
```